---
title: "Darwin Core mapping"
subtitle: "For: Monitoring of invasive alien species by the Provincial Center of Environmental Research, Province East-Flanders, Belgium"
author:
- Damiano Oldoni
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(readr)          # To read data
library(dplyr)          # To do data wrangling
library(tidylog)        # To provide feedback on dplyr functions
library(here)           # To find files
library(sf)             # To convert coordinate systems
library(DBI)            # To work with databases
library(RSQLite)        # To work with SQLite databases in R
```

## Read source data

Create a data frame `occurrences` from the source data:

```{r read_inputs}
occurrences <- readr::read_csv(
  file = here::here("data", "raw", "invasieve_uitheemse_soorten.csv"),
  col_types = readr::cols(.default = readr::col_character())
)
```

Preview data:

```{r preview_input_data}
occurrences %>% head(n = 5)
```

# Change coordinate system

Data are provided in Belgian Lambert72 ([EPSG 31370](https://epsg.io/31370)) coordinates. DwC accepts data in WGS84 ([EPSG 4326](https://epsg.io/4326)) only.

```{r}
occurrences <- 
  occurrences %>% 
  st_as_sf(crs = st_crs(31370), wkt = "Shape") %>%
  st_transform(crs = 4326)
coords <- dplyr::as_tibble(st_coordinates(occurrences))
occurrences <- 
  dplyr::as_tibble(occurrences) %>% 
  dplyr::bind_cols(coords) %>%
  dplyr::select(-.data$Shape)
```

Preview:

```{r}
occurrences %>% head(5)
```

# Create database

Create a SQLite database with the source data, so it can be queried with SQL in the next steps:

```{r create_db}
message("Create in memory SQLite database...")
con <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
# import occurrences
DBI::dbWriteTable(con, "occurrences", occurrences)
message("DONE")
```

# Darwin Core mapping

Create [Occurrence](https://rs.gbif.org/core/dwc_occurrence_2022-02-02.xml) extension:

```{r occurrence}
message("Map occurrences to DwC...")
dwc_occurrence_sql <- glue::glue_sql(
  readr::read_file(here::here("sql", "dwc_occurrence.sql")), 
  .con = con)
dwc_occurrence <- DBI::dbGetQuery(con, dwc_occurrence_sql)
message("DONE")
```

# Save data to CSV

```{r save_csv}
readr::write_csv(
  dwc_occurrence,
  here::here("data", "processed", "occurrence.csv"),
  na = ""
)
```

# Overview changes

In this section we summarize the changes in the DwC output after the very last udpate.

## Read DwC files

Read DwC mapping files from both `main`:

```{r main}
dwc_occurrence_main <- readr::read_csv(
  file = "https://raw.githubusercontent.com/riparias/pov-ias-plants-occurrences/main/data/processed/occurrence.csv",
  col_types = readr::cols(.default = "c"))
```

and actual branch (`automatic-update-*`):
 
```{r automatic-update}
dwc_occurrence_update <- readr::read_csv(
  file = here::here("data","processed","occurrence.csv"),
  col_types = readr::cols(.default = "c"))
```

## Overview

How many new occurrences, i.e. new `occurrenceID` values, have been added?

```{r new_occurrenceID_values}
new_occs <- 
  dwc_occurrence_update %>%
  filter(!.data$occurrenceID %in% dwc_occurrence_main$occurrenceID)
message("New occurrences:")
print(new_occs)
```

Have some occurrences been removed?

```{r removed_occs}
removed_occs <- 
  dwc_occurrence_main %>%
  filter(!.data$occurrenceID %in% dwc_occurrence_update$occurrenceID)
message("Removed occurrences:")
print(removed_occs)
```

Total number of rows of new DwC output in comparison with the previous version:

```{r nrows}
message(paste("Number of occurrences (new):", nrow(dwc_occurrence_update)))
message(paste("Number of occurrences (old):", nrow(dwc_occurrence_main)))
message(paste("Difference:", 
              nrow(dwc_occurrence_update) - nrow(dwc_occurrence_main))
)
```

## New `ScientificName` values

The DwC `scientificName` is a manual mapping of scientific names from field `ScientificName`. It is therefore important to get a list with all new values:

```{r new_ScientificName}
occs_without_sc_name <- 
  dwc_occurrence_update %>%
  filter(is.na(scientificName)) %>%
  pull(occurrenceID)
query <- glue::glue_sql("
  SELECT 
    DISTINCT o.\"ScientificName\"
  FROM occurrences AS o 
  WHERE  
  o.\"ObservationIdentifier\" IN ({occs_without_sc_name*})
  ", .con = con)
query
raw_sc_names <- dplyr::as_tibble(DBI::dbGetQuery(con, query))
message("New values in field 'ScientificName' to be mapped:")
print(raw_sc_names)
```
